# Libraries
```{r}
library(here)
library(tidyverse)
library(haven)
```

# Check1: Stata match with NPPES (both entity type) and CMHC NPI from Find_NPI.R
Are the CMHC NPI associated with entity type = 1 happening at the same time as an org type CMHC
```{r}
indiv_cmhc <- read_dta(here("Trunk", "Derived", "NBER_NPPES", "known_cmhc.dta")) %>% 
  filter(entity == 1) %>% 
  pull(NPI)

# Check with IL: are there same-day claims?
# Claims that are billed by the individual type NPI (that was at least once coupled with POS_CD = CMHC)
indiv_dates <- open_dataset("/gpfs/milgram/pi/medicaid_lab/data/cms/ingested/TMSIS_taf/taf_other_services_header/year=2022/state=IL/data.parquet") |>
  select(BENE_ID, CLM_ID, BLG_PRVDR_NPI, SRVC_BGN_DT) |>
  filter(BLG_PRVDR_NPI %in% indiv_cmhc) |>
  collect() |>
  filter(!is.na(BENE_ID)) |>
  rename(indiv_claim = CLM_ID,
         indiv_NPI = BLG_PRVDR_NPI)

# Find the claims with the same beneficiary on the same date 
same_day_visit <- open_dataset("/gpfs/milgram/pi/medicaid_lab/data/cms/ingested/TMSIS_taf/taf_other_services_header/year=2022/state=IL/data.parquet") |>
  select(BENE_ID, CLM_ID, BLG_PRVDR_NPI, SRVC_BGN_DT, POS_CD) |>
  filter(!CLM_ID %in% indiv_dates$indiv_claim) |>
  inner_join(indiv_dates, by = c("BENE_ID", "SRVC_BGN_DT")) |>
  collect()

# Of all the claims that were on the same day, remove if there is no chance of being a CMHC 
# (somewhere else is already specified by the POS_CD)
cmhc_or_na <- same_day_visit %>% 
  filter(POS_CD == "53" | is.na(POS_CD)) %>% 
  mutate(match = 1) %>% 
  select(indiv_claim, POS_CD, match) %>% 
  right_join(indiv_dates, by = "indiv_claim") %>% 
  group_by(indiv_claim) %>% 
  slice(1) %>% 
  mutate(match = if_else(is.na(match), 0, match))

# Of the matches, how many are CMHCs for sure (through POS) and how many are NA?
na_summary <- cmhc_or_na %>% 
  filter(match == 1) %>% 
  group_by(POS_CD) %>% 
  summarise(count = n())
```

# Check2: What are the non-matches?
There are 6% of non-matches, even when I include the individual-type NPI. Compare the number 
of claims that are associated with non-matches, organization-type NPI, and individual-type NPI
```{r}
indiv_npi <- read_dta(here("Trunk", "Derived", "NBER_NPPES", "known_cmhc.dta")) %>% 
  filter(entity == 1) %>% 
  pull(NPI)

org_npi <- read_dta(here("Trunk", "Derived", "NBER_NPPES", "known_cmhc.dta")) %>% 
  filter(entity == 2) %>% 
  pull(NPI)

nonmatch <- open_dataset(here("Trunk", "Derived", "all_states_cmhc.parquet")) |>
  collect() |>
  filter(!is.na(BLG_PRVDR_NPI)) |>
  filter(!(BLG_PRVDR_NPI %in% indiv_npi | BLG_PRVDR_NPI %in% org_npi)) |>
  rename(NPI = BLG_PRVDR_NPI) |>
  distinct(NPI) |>
  pull(NPI)

# Check with IL 
claim_count <- open_dataset("/gpfs/milgram/pi/medicaid_lab/data/cms/ingested/TMSIS_taf/taf_other_services_header/year=2022/state=IL/data.parquet") |>
  filter(BLG_PRVDR_NPI %in% indiv_npi | BLG_PRVDR_NPI %in% org_npi | BLG_PRVDR_NPI %in% nonmatch) |>
  select(CLM_ID, BENE_ID, SRVC_BGN_DT, BLG_PRVDR_NPI) |>
  collect() |>
  mutate(npi_type = case_when(BLG_PRVDR_NPI %in% indiv_npi ~ "individual",
                              BLG_PRVDR_NPI %in% org_npi ~ "organization",
                              BLG_PRVDR_NPI %in% nonmatch ~ "nonmatch")) |>
  # If it's same-day same-person, overwrite 
  # organization overwrites both, individual overwrites nonmatch 
  group_by(BENE_ID, SRVC_BGN_DT) |>
  mutate(npi_type = case_when(any(npi_type == "organization") ~ "organization",
                              any(npi_type == "individual") ~ "individual",
                              TRUE ~ "nonmatch")) |>
  slice(1) |>
  ungroup() |>
  group_by(npi_type) |>
  summarise(count = n_distinct(CLM_ID)) |>
  # Make proportions 
  mutate(total = sum(count),
         proportion = round(count/total*100, 2)) 
  
  
```


