Making new patient characteristic table (table 1), comparing between repeater type 

# Libraries
```{r}
library(tidyverse)
library(here)
```

# Data 
```{r}
# First trying with Illinois only 
states <- c("IL")

all_states <- vector("list", length = length(states))
for(st in seq_along(states)){
  
  save_name <- paste0("tab1_chars_", states[[st]], ".csv")
  state <- read_csv(here("Trunk", "Derived", "new_patient_char", save_name))
  
  all_states[[st]] <- state
}

all_states <- bind_rows(all_states)
```

# Lists 
```{r}
# Diagnosis SMI 
DGNS_CD_SCZ <- c('F200', 'F201', 'F202', 'F203', 'F205', 'F2081', 'F2089', 'F209', 'F250', 'F251', 'F258', 'F259')

DGNS_CD_BD <- c('F3010','F3011','F3012','F3013','F302','F303','F304','F308','F309','F310','F3110','F3111','F3112','F3113','F312','F3130','F3131','F3132','F314','F315','F3160','F3161','F3162','F3163','F3164','F3170','F3171','F3172','F3173','F3174','F3175','F3176','F3177','F3178','F3181','F3189','F319')

DGNS_CD_MDD <- c('F320','F321','F322','F323','F324','F325','F329','F330','F331','F332','F333','F3340','F3341','F3342','F339','F341')

DGNS_CD_PSYCH <- c('F060', 'F061', 'F062', 'F0630', 'F0631', 'F0632', 'F0633', 'F0634', 'F064', 'F0670', 'F0671', 'F068', 'F22', 'F23', 'F24', 'F28', 'F29', 'F302', 'F312', 'F315', 'F3164', 'F323', 'F333', 'F4489')

smi <- c(DGNS_CD_BD, DGNS_CD_MDD, DGNS_CD_SCZ, DGNS_CD_PSYCH)

code_zero <- grep("0$", smi, value = TRUE)
code_nozero <- sub("0$", "", code_zero)

smi_code <- unique(c(smi, code_nozero))

smi_code <- gsub("\\.", "", smi_code)

# Diagnosis SUD 
# ICD-10 codes F10-F19 (this is the same way the SUD_DGNS_IND is coded in TAF)
sud_list <- paste0("F", 10:19)

```


# Making table 1 (demographics and previous health history for new patients)
```{r}
check <- all_states %>% 
  # Check for the first visit diagnosis type 
  mutate(visit1_SUD = if_else(substr(visit1_dgns_1, 1, 3) %in% sud_list | substr(visit1_dgns_2, 1, 3) %in% sud_list, 1, 0),
         visit1_SMI = if_else(visit1_dgns_1 %in% smi_code | visit1_dgns_2 %in% smi_code, 1, 0)) %>% 
  mutate(
    spell_length = ifelse(repeat_visits == 0, "One Visit", "More Than One Visit"),
    spell_length = factor(spell_length, levels = c("One Visit", "More Than One Visit")),

    # Age groups
    age_group = case_when(
      AGE >= 13 & AGE <= 17 ~ "13-17",
      AGE >= 18 & AGE <= 25 ~ "18-25",
      AGE >= 26 & AGE <= 45 ~ "26-45",
      AGE >= 46 & AGE <= 62 ~ "46-62",
      TRUE ~ NA_character_
    ),
    age_group = factor(age_group, levels = c("13-17", "18-25", "26-45", "46-62")),

    # Sex
    sex_label = ifelse(SEX_CD == "F", "Female", "Male"),

    # Binary flags — adjust the values if yours differ
    visit1_SMI = as.numeric(visit1_SMI),
    visit1_SUD = as.numeric(visit1_SUD),
    last_year_bh_ip  = as.numeric(last_year_bh_ip),
    last_year_bh_ed  = as.numeric(last_year_bh_ed),
    last_year_any_ed = as.numeric(last_year_any_ed)
  )

# Create functions 
# For catergorical variables
cat_row <- function(data, var, level, label){
  data %>% 
    group_by(spell_length) %>% 
    summarise(n = sum(.data[[var]] == level, na.rm = TRUE),
              pct = mean(.data[[var]] == level, na.rm = TRUE) * 100,
              .groups = "drop") %>% 
    mutate(cell = sprintf("%d (%.1f%%)", n, pct)) %>%
    select(spell_length, cell) %>%
    tidyr::pivot_wider(names_from = spell_length, values_from = cell) %>%
    mutate(Category = "", Variable = label) %>%
    select(Category, Variable, `One Visit`, `More Than One Visit`)
}

# For binary variables 
bin_row <- function(data, var, label) {
  data %>%
    group_by(spell_length) %>%
    summarise(
      n   = sum(.data[[var]] == 1, na.rm = TRUE),
      pct = mean(.data[[var]] == 1, na.rm = TRUE) * 100,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%d (%.1f%%)", n, pct)) %>%
    select(spell_length, cell) %>%
    tidyr::pivot_wider(names_from = spell_length, values_from = cell) %>%
    mutate(Category = "", Variable = label) %>%
    select(Category, Variable, `One Visit`, `More Than One Visit`)
}

# This is done differently because the denominator of the percentage is different 
# number of patients with one visit / total number of new patients 
total_row <- check %>%
  group_by(spell_length) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(
    pct  = n / sum(n) * 100,
    cell = sprintf("%d (%.1f%%)", n, pct)
  ) %>%
  select(spell_length, cell) %>%
  tidyr::pivot_wider(names_from = spell_length, values_from = cell) %>%
  mutate(Category = "Total (N(%))", Variable = "") %>%
  select(Category, Variable, `One Visit`, `More Than One Visit`)


age_rows <- bind_rows(
  cat_row(check, "age_group", "13-17", "13-17"),
  cat_row(check, "age_group", "18-25", "18-25"),
  cat_row(check, "age_group", "26-45", "26-45"),
  cat_row(check, "age_group", "46-62", "46-62")
) %>% mutate(Category = ifelse(row_number() == 1, "Age in 2021", ""))

sex_rows <- bind_rows(
  cat_row(check, "sex_label", "Female", "Female"),
  cat_row(check, "sex_label", "Male",   "Male")
) %>% mutate(Category = ifelse(row_number() == 1, "Sex", ""))

dx_rows <- bind_rows(
  bin_row(check, "visit1_SMI", "Any SMI"),
  bin_row(check, "visit1_SUD", "Any SUD")
) %>% mutate(Category = ifelse(row_number() == 1, "Diagnosis at first visit", ""))

hx_rows <- bind_rows(
  bin_row(check, "last_year_bh_ip",  "Behavioral Health Related Inpatient Care"),
  bin_row(check, "last_year_bh_ed",  "Behavioral Health Related ED Visit"),
  bin_row(check, "last_year_any_ed", "Any ED Visit")
) %>% mutate(Category = ifelse(row_number() == 1, "Previous year health history", ""))

table_data <- bind_rows(total_row, age_rows, sex_rows, dx_rows, hx_rows)


```

# Repeat number histogram
```{r}
library(ggplot2)

# Only look at people with SMI diagnosis 
all_states <- all_states %>% 
  filter(visit1_dgns_1 %in% smi_code | visit1_dgns_2 %in% smi_code)

ggplot(all_states, aes(x = repeat_visits)) + 
  geom_histogram(binwidth = 20, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  labs(x = "Number of visits")


suppress_20 <- all_states %>% 
  mutate(repeat_visits = if_else(repeat_visits > 20, 20, repeat_visits))

ggplot(suppress_20, aes(x = repeat_visits)) + 
  geom_histogram(binwidth = 1, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  labs(x = "Number of visits")


```

# Making table 3 ver1 (post-CMHC BH envounters for new patients)
```{r}
all_states <- read_csv(here("Trunk", "Derived", "new_patient_char", "post_cmhc", "alive_post_cmhc.csv"))

# For binary variables 
bin_row <- function(data, var, label) {
  data %>%
    group_by(spell_length) %>%
    summarise(
      n   = sum(.data[[var]] == 1, na.rm = TRUE),
      pct = mean(.data[[var]] == 1, na.rm = TRUE) * 100,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%d (%.1f%%)", n, pct)) %>%
    select(spell_length, cell) %>%
    tidyr::pivot_wider(names_from = spell_length, values_from = cell) %>%
    mutate(Category = "", Variable = label) %>%
    select(Category, Variable, `One Visit`, `More Than One Visit`)
}

tab3_dt <- all_states %>% 
  # Count people without any BH visit instead
  mutate(none_bh = if_else(any_bh_visit == 1, 0, 1)) %>% 
  mutate(spell_length = ifelse(repeat_visits == 0, "One Visit", "More Than One Visit"),
          spell_length = factor(spell_length, levels = c("One Visit", "More Than One Visit")))

tab3_rows <- bind_rows(
  bin_row(tab3_dt, "crisis_services",  "Crisis Services"),
  bin_row(tab3_dt, "mh_other_facility",  "MH tx at non-CMHCi"),
  bin_row(tab3_dt, "mh_other_cmhc", "MH tx at different CMHC"),
  bin_row(tab3_dt, "none_bh", "No outpatient BH visit"),
  bin_row(tab3_dt, "bh_ed_visit", "Any ED BH visit"),
  bin_row(tab3_dt, "bh_inpatient", "Any IP BH visit"),
  bin_row(tab3_dt, "DEATH_IND", "Death")
) %>% 
  select(-c("Category"))
```


# Making table 3 ver 2 (same content, row by MH service presence and location, col by death, ED, IP)
```{r}
# Treat death completely separately 
death <- read_csv(here("Trunk", "Derived", "new_patient_char", "post_cmhc", "dead_post_cmhc.csv")) 

# Define a reusable summary function for one group's data
summarise_mh <- function(df) {
  tibble(
    row_label = c(
      "Total (N)",
      "Any MH Tx - At a different CMHC",
      "Any MH Tx - At a non-CMHC",
      "Total any MH (not CMHCi)",
      "No MH Tx after last CMHC visit"
    ),
    n = c(
      nrow(df),
      sum(df$mh_other_cmhc == 1, na.rm = TRUE),
      sum(df$mh_non_cmhc   == 1, na.rm = TRUE),
      sum(df$any_mental    == 1, na.rm = TRUE),
      sum(df$any_mental    == 0, na.rm = TRUE)
    )
  )
}

# Apply across groups, then pivot so each group becomes a column
result_pct <- death %>%
  group_by(one_time_visit) %>%
  group_modify(~ summarise_mh(.x)) %>%
  ungroup() %>%
  group_by(one_time_visit) %>%
  mutate(
    total = n[row_label == "Total (N)"],
    n_pct = paste0(n, " (", round(n / total * 100, 1), "%)")
  ) %>%
  ungroup() %>%
  select(-n, -total) %>%
  pivot_wider(
    names_from  = one_time_visit,
    values_from = n_pct
  )


# Alive people 
alive <- read_csv(here("Trunk", "Derived", "new_patient_char", "post_cmhc", "alive_post_cmhc.csv"))

all_patients <- alive %>%
  group_by(one_time_visit) %>%
  group_modify(~ summarise_mh(.x)) %>%
  ungroup() %>%
  group_by(one_time_visit) %>%
  mutate(
    total = n[row_label == "Total (N)"],
    n_pct = paste0(n, " (", round(n / total * 100, 1), "%)")
  ) %>%
  ungroup() %>%
  select(-n, -total) %>%
  pivot_wider(
    names_from  = one_time_visit,
    values_from = n_pct
  )

had_ed <- alive %>% 
  filter(bh_ed_visit == 1) %>% 
  group_by(one_time_visit) %>%
  group_modify(~ summarise_mh(.x)) %>%
  ungroup() %>%
  group_by(one_time_visit) %>%
  mutate(
    total = n[row_label == "Total (N)"],
    n_pct = paste0(n, " (", round(n / total * 100, 1), "%)")
  ) %>%
  ungroup() %>%
  select(-n, -total) %>%
  pivot_wider(
    names_from  = one_time_visit,
    values_from = n_pct
  )

had_ip <- alive %>% 
  filter(bh_inpatient == 1) %>% 
  group_by(one_time_visit) %>%
  group_modify(~ summarise_mh(.x)) %>%
  ungroup() %>%
  group_by(one_time_visit) %>%
  mutate(
    total = n[row_label == "Total (N)"],
    n_pct = paste0(n, " (", round(n / total * 100, 1), "%)")
  ) %>%
  ungroup() %>%
  select(-n, -total) %>%
  pivot_wider(
    names_from  = one_time_visit,
    values_from = n_pct
  )

# Make the table 
one_visit <- cbind(all_patients[,1:2], had_ed[,2], had_ip[,2]) 

colnames(one_visit) <- c("MH", "All patients","Had BH ED visit", "Had BH inpatient")


```

# Table 3 ver 3 (column wise percentage -> rowwise percentage)
```{r}
# ----------------------------------------------------------
# 0. Combine datasets
# ----------------------------------------------------------
alive <- read_csv(here("Trunk", "Derived", "new_patient_char", "post_cmhc", "alive_post_cmhc.csv"))
dead <- read_csv(here("Trunk", "Derived", "new_patient_char", "post_cmhc", "dead_post_cmhc.csv")) 

alive$died <- 0
dead$died  <- 1

df <- bind_rows(alive, dead)

# ----------------------------------------------------------
# 1. Helper: build one panel's rows
# ----------------------------------------------------------
make_panel <- function(data) {

  total_n <- nrow(data)

  # --- compute row-level flags (mutually exclusive groups) ---
  # "Any MH Tx" rows: any_mental == 1
  # "No MH Tx"  row:  any_mental == 0

  row_stats <- function(subset_mask, label1, label2 = NULL) {
    sub   <- data[subset_mask, ]
    n_sub <- nrow(sub)

    baseline_pct <- sprintf("%d (%.1f%%)", n_sub, 100 * n_sub / total_n)

    # Had any ED BH visit – row percent within the subset
    ed_n   <- sum(sub$bh_ed_visit,  na.rm = TRUE)          # <-- adjust var name if needed
    ip_n   <- sum(sub$bh_inpatient,  na.rm = TRUE)          # <-- adjust var name if needed
    died_n <- sum(sub$died,       na.rm = TRUE)

    ed_pct   <- if (n_sub > 0) sprintf("%d (%.1f%%)", ed_n,   100 * ed_n   / n_sub) else "—"
    ip_pct   <- if (n_sub > 0) sprintf("%d (%.1f%%)", ip_n,   100 * ip_n   / n_sub) else "—"
    died_pct <- if (n_sub > 0) sprintf("%d (%.1f%%)", died_n, 100 * died_n / n_sub) else "—"

    data.frame(
      label1       = label1,
      label2       = if (!is.null(label2)) label2 else "",
      baseline     = baseline_pct,
      ed_bh        = ed_pct,
      ip_bh        = ip_pct,
      died         = died_pct,
      stringsAsFactors = FALSE
    )
  }

  # Total row
  total_row <- data.frame(
    label1   = "Total",
    label2   = "",
    baseline = sprintf("%d (100%%)", total_n),
    ed_bh    = sprintf("%d (%.1f%%)", sum(data$bh_ed_visit, na.rm=TRUE),
                       100 * mean(data$bh_ed_visit, na.rm=TRUE)),
    ip_bh    = sprintf("%d (%.1f%%)", sum(data$bh_inpatient, na.rm=TRUE),
                       100 * mean(data$bh_inpatient, na.rm=TRUE)),
    died     = sprintf("%d (%.1f%%)", sum(data$died, na.rm=TRUE),
                       100 * mean(data$died, na.rm = TRUE)),
    stringsAsFactors = FALSE
  )

  # Any MH Tx rows
  r1 <- row_stats(data$any_mental == 1 & data$mh_other_cmhc == 1,
                  "Any MH Tx after last CMHC visit", "At a different CMHC")
  r2 <- row_stats(data$any_mental == 1 & data$mh_non_cmhc == 1,
                  "", "At a non-CMHC")
  r3 <- row_stats(data$any_mental == 1,
                  "", "Total any MH (not CMHCi)")

  # No MH Tx row
  r4 <- row_stats(data$any_mental == 0,
                  "No MH Tx after last CMHC visit", "")

  bind_rows(total_row, r1, r2, r3, r4)
}

# ----------------------------------------------------------
# 2. Split by panel
# ----------------------------------------------------------
panel_a_data <- df %>% filter(one_time_visit == 1)
panel_b_data <- df %>% filter(one_time_visit == 0)

n_a     <- nrow(panel_a_data)
n_total <- nrow(df)
pct_a   <- round(100 * n_a / n_total, 1)

n_b   <- nrow(panel_b_data)
pct_b <- round(100 * n_b / n_total, 1)

panel_a <- make_panel(panel_a_data)
panel_b <- make_panel(panel_b_data)

# ----------------------------------------------------------
# 3. Add panel header rows and combine
# ----------------------------------------------------------
header_row <- function(label) {
  data.frame(label1=label, label2="", baseline="", ed_bh="", ip_bh="", died="",
             stringsAsFactors=FALSE)
}

table_out <- bind_rows(
  header_row(sprintf("PANEL A: PATIENTS WITH ONE CMHC VISIT (N = %d, %.1f%%)", n_a, pct_a)),
  panel_a,
  header_row(sprintf("PANEL B: PATIENTS WITH MORE THAN ONE CMHC VISIT (N = %d, %.1f%%)", n_b, pct_b)),
  panel_b
)

# ----------------------------------------------------------
# 4. Clean up column names for display
# ----------------------------------------------------------
names(table_out) <- c(
  "Row Group",
  "Sub-row",
  "All at baseline (death not included) [column %]",
  "Had any ED BH visit [row %]",
  "Had any IP BH visit [row %]",
  "Died [row %]"
)

# ----------------------------------------------------------
# 5. Print to console
# ----------------------------------------------------------
print(table_out, row.names = FALSE)

write.csv(table_out, here("Output", "Tables", "tab_3_ver3.csv"), row.names = FALSE)
```

